---
title: "logistic_regression?"
format: html
editor: visual
---

```{r}
suppressMessages(library(tidyverse))
theme_set(theme_bw())

drinking_water <- read_csv(here::here('data', "drinking_water.csv"), show_col_types = FALSE)
```

# Questions

- What do you think each row represents?

a observation at a given year

- What columns do you think represent:
    - Drinking water health violations?
      'health'
      
    - Percent Black and Hispanic population in the utility district?
      'pctblack', 'pcthisp'  
    
    - Median household income in the utility district?
      'medianincomehouse'
    
    
- One column includes a count of drinking water health violations. How would you create a new column with a binary variable representing whether there were any violations?
```{r}
# mutate
# ifelse vs if_else
# ifelse = base, if_else = tidyverse
# base will let you combine data types

drinking_water <- drinking_water %>% 
  mutate(violation = if_else(health > 0, 1, 0))
```
  
    
- Create a scatter plot of violations against race (percent Black population), ethnicity (percent Hispanic population), and SES (median household income). What visualization issues do we get with a scatter plot? How could you address that?

The points are overplotted, so we can summarize within bins

```{r}
drinking_water_binned <- drinking_water %>%
  mutate(pcthisp = round(pcthisp * 5) / 5) %>%
  group_by(pcthisp) %>%
  summarize(violation = mean(violation))

ggplot(drinking_water_binned, aes(pcthisp, violation)) + geom_point()
```

```{r}
pcthisp_lm <- lm(violation ~ pcthisp, drinking_water_binned)
summary(pcthisp_lm)
```

## Questions

- Plot the residuals. What pattern do you notice?
```{r}
resids <- as.tibble(resid(pcthisp_lm))

drinking_water_complete <- drop_na(drinking_water_binned, pcthisp, violation)

drinking_water_complete <- drinking_water_complete %>% mutate(resids = resid(pcthisp_lm))

ggplot(drinking_water_complete, aes(resids, violation)) + 
  geom_point()
```


### Make a plot of the residuals.

- Plot the raw data and the predicted probabilities for violations. Are there any obvious problems?

### Plot raw data and predicted probabilities

```{r}
ggplot(drinking_water_binned, aes(pcthisp, violation)) + 
  geom_point() +
  geom_smooth(method = lm)
```

```{r}
# Inverse logit utility function
inv_logit <- function(x) exp(x) / (1 + exp(x))
  
# Likelihood of the coefficients, given the data
likelihood_fun <- function(coefficients, data) {
  # coefficients looks like: c(beta0 = ???, beta1 = ???)
  # Calculate logit(p) based on coefficients and predictor
  logit_p <- coefficients['beta0'] + coefficients['beta1'] * data$pcthisp

  # Invert the logit to get p
  p <- inv_logit(logit_p)
  
  # Use the PMF of the Bernoulli to get our log likelihoods
  loglik <- dbinom(data$violation, size = 1, prob = p, log = TRUE)
  
  # Sum the negative log likelihood
  negloglik <- -sum(loglik)
}

# Use an optimization function to get the maximum likelihood coefficients
drinking_water_complete <- drop_na(drinking_water, pcthisp, violation)
coef_optim <- optim(c(beta0 = 0, beta1 = 0), 
                    likelihood_fun, 
                    data = drinking_water_complete)
```

What were your maximum likelihood estimates for and...


# What are the maximum likelihood estimates for our coefficients?
```{r}
coef_optim$par
beta0_hat <- coef_optim$par['beta0']
beta1_hat <- coef_optim$par['beta1'] 
```


# Hint: explore coef_optim

    What’s the predicted probability of drinking water violations for communities with 0%, 50%, and 100% Hispanic population?
    Plot the predicted probability across the whole range 0-100% Hispanic.

# Create and plot predictions

- How much does the probability of a drinking water violation change when percent Hispanic population increases from 10 to 20%, 45 to 55%, and 80 to 90%?

```{r}
logit_p <- beta0_hat + beta1_hat * c(0,50,100)
p <- inv_logit(logit_p)
p

# create and plot predictions
tibble(pcthisp = 0:100,
       logit_p = beta0_hat + beta_1_hat * pcthisp,
       p = inv_logit(logit_p)) %>% 
  ggplot(aes(pcthisp, p)) +
  geom_line()


```



    How would you interpret the coefficients? What do the slope and intercept mean in this context? Where is the relationship linear, and where is it non-linear?
    Create a “DEM” of the likelihood landscape for 

and . Choose a range of and values around your best estimates, calculate the likelihood for each combination, and create a figure with on the x-axis, on the y-axis, and the likelihood as the fill. Add a point for
.
Bonus problem: add contours!